% Data


```{r setup, include=FALSE}
library(readxl)
library(tidyr)
library(dplyr)
library(tibble)
library(ggplot2)
```


## Read data 
 
The raw data were organized in excel sheets. Let's load all sheets using the `read_excel` function and bind using `bind_rows` to make a single data set.

```{r}
path <- "data/Summary of raw data for CLR Epidemic 1.xlsx"

data <- read_excel(path, sheet = "Agaro") %>% 
  bind_rows(
    read_excel(path, sheet = "Gera"),
    read_excel(path, sheet = "Jimma"),
    read_excel(path, sheet = "Metu"),
    read_excel(path, sheet = "Tepi")
  )

```

### Data cleaning

- The data from Jimma got some empty cells in the `Year` column that need to be removed. 
- Some variety names are duplicate, in some manner, that needs to be fixed. 

```{r}
CLR <- data %>%
  filter(!is.na(year)) %>% 
  mutate(variety = recode(variety, "7475" = "7454", "741140" = "74140", "f-35" = "F-35", "f-59" = "F-59"))
```

### Store data as .csv file

```{r}
write.csv(CLR, "data/CLR.csv")
```

## AUDPC calculation

### CLR Severity

```{r}
library(agricolae)

Sev <- CLR %>%
    mutate(month = recode(month, `8` = 1, `9` = 2, `10` = 3, `11` = 4, `12` = 5, `1` = 6, `2` = 7, `3` = 8, `4` = 9, `5` = 10, `6` = 11, `7` = 12)) %>%
  select(year, site, variety, tree, branch, month, severity) %>%
  group_by(year, site, variety, tree, branch, month) %>%
  mutate(rn = row_number()) %>% 
  ungroup %>%
  spread(month, severity) %>%
  mutate_if(is.numeric, ~replace_na(., 0))

ratings<- Sev[,c(7:18)]

#For Months
months<-c(1,2,3,4,5,6,7,8,9,10,11,12)
Sev$AUDPCm <- audpc(ratings, months, type="absolute")

# For days
days<-c(1,32,62,93,123,154,185,213,244,274,305,335)
Sev$AUDPCd <- audpc(ratings, days, type="absolute")

# Unlist AUDPC values
Sev$AUDPCm <- as.numeric(unlist(Sev$AUDPCm))
```

### Store data as .csv file

```{r}
Sev %>% 
  select(-AUDPCd) %>%
  write.csv("data/clr_audpc.csv")
```

