% Mixed Models


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(tidyverse)
library(cowplot)
library(ggthemes)
library(gt)
theme_set(theme_minimal_grid())

```

# Setup

## Read data

```{r}
dat <- read_csv("data/clr_audpc.csv")
```

## Prepare variables

```{r}
# Calculate max. severity column
dat1 <- dat %>% 
  group_by(X1) %>% 
  mutate(max = max(`1`,`2`,`3`, `4`, `5`, `6`, `7`, `8`, `9`, `10`, `11`, `12`))

# Calculate mean AUDPC and Max. severity  
sev <- dat1 %>% 
  group_by(year, site, variety, tree) %>% 
  summarize(mean_max = mean(max), mean_audpc = mean(AUDPCm))  
```


# Data exploration

## Relationship between AUDPC and disease severity

```{r}
sev %>%  
    ggplot(aes(log(mean_max), log(mean_audpc)))+
  geom_point()
```

# Data analysis

## Agaro 

```{r}
sev_agaro <- sev %>% 
  filter(site == "agaro")
```

### Boxplot

```{r}
sev_agaro %>% 
  ggplot(aes(reorder(variety, mean_audpc), mean_audpc))+
  geom_boxplot()
```

### AUDPC vs Max. severity

```{r}
sev_agaro %>% 
  ggplot(aes(mean_audpc, mean_max, color = variety))+
  geom_point()
```


### Mixed model fit

**Fit and residuals**

```{r}
library(lme4)
m_audpc <- lmer(mean_audpc ~ variety + (1 | year), data = sev_agaro)

library(DHARMa)
simres <- simulateResiduals(m_audpc)
plot(simres)
```

**Obtain estimates of the means and perform multiple comparison**

### Scott-Knott

```{r}
library(ScottKnott)

sev_audpc_agaro_sk <- with(sev_agaro,
              SK(x= variety,
                 y=mean_audpc,
                 model='y ~ x',
                 which='x'))

plot(sev_audpc_agaro_sk)
```

### Mean AUDPC

```{r}
sev_agaro_sk <- summary(sev_audpc_agaro_sk)
```


### Means and boxplot

```{r}
library(patchwork)

# severity
p2 <- sev_agaro %>% 
  ggplot(aes(reorder(variety, mean_audpc), mean_audpc))+
  geom_boxplot()+
  labs(x = "", y = "AUDPC", title = "Site: Agaro", subtitle = "Severity")+
  coord_flip()

sev_agaro_sk$Means <- round(sev_agaro_sk$Means, 1)

# Means table
library(gridExtra)
p2 + gridExtra::tableGrob(sev_agaro_sk %>% rename(Cultivar = Levels), rows = rownames("Levels"), 
                          theme = ttheme_minimal(padding = unit(c(4, 4.5), "mm")))
```

## Gera

```{r}
sev_gera <- sev %>% 
  filter(site == "gera")
```

### Boxplot

```{r}
sev_gera %>% 
  ggplot(aes(reorder(variety, mean_audpc), mean_audpc))+
  geom_boxplot()
```

### AUDPC vs Max. severity

```{r}
sev_gera %>% 
  ggplot(aes(mean_audpc, mean_max, color = variety))+
  geom_point()
```

### Mixed model fit

**Fit and residuals**

```{r}
m_audpc <- lmer(mean_audpc ~ variety + (1 | year), data = sev_gera)
simres <- simulateResiduals(m_audpc)
plot(simres)
```

**Obtain estimates of the means and perform multiple comparison**

### Scott-Knott

```{r fig.height=6, fig.width=12}
sev_audpc_gera_sk <- with(sev_gera,
              SK(x= variety,
                 y=mean_audpc,
                 model='y ~ x',
                 which='x'))

plot(sev_audpc_gera_sk)
```

### Mean AUDPC

```{r}
sev_gera_sk <- summary(sev_audpc_gera_sk)
```

### Means and boxplot

```{r fig.height=7.5, fig.width=6}
# severity
p2 <- sev_gera %>% 
  ggplot(aes(reorder(variety, mean_audpc), mean_audpc))+
  geom_boxplot()+
  labs(x = "", y = "AUDPC", title = "Site: Gera", subtitle = "Severity")+
  coord_flip()

sev_gera_sk$Means <- round(sev_gera_sk$Means, 1)

# Means table
p2 + gridExtra::tableGrob(sev_gera_sk %>% rename(Cultivar = Levels), rows = rownames("Levels"), 
                          theme = ttheme_minimal(padding = unit(c(4, 4.5), "mm")))
```

## Jimma

```{r}
sev_jimma <- sev %>% 
  filter(site == "jimma")
```

### Boxplot

```{r}
sev_jimma %>% 
  ggplot(aes(reorder(variety, mean_audpc), mean_audpc))+
  geom_boxplot()
```

### AUDPC vs Max. severity

```{r}
sev_jimma %>% 
  ggplot(aes(mean_audpc, mean_max, color = variety))+
  geom_point()
```

### Mixed model fit

**Fit and residuals**

```{r}
m_audpc <- lmer(mean_audpc ~ variety + (1 | year), data = sev_jimma)
simres <- simulateResiduals(m_audpc)
plot(simres)
```

**Obtain estimates of the means and perform multiple comparison**

### Scott-Knott

```{r fig.height=5, fig.width=10}
sev_audpc_jimma_sk <- with(sev_jimma,
              SK(x= variety,
                 y=mean_audpc,
                 model='y ~ x',
                 which='x'))

plot(sev_audpc_jimma_sk)
```


### Means

```{r}
sev_jimma_sk <- summary(sev_audpc_jimma_sk)
```

### Means and boxplot

```{r fig.height=6, fig.width=6}
# severity
p2 <- sev_jimma %>% 
  ggplot(aes(reorder(variety, mean_audpc), mean_audpc))+
  geom_boxplot()+
  labs(x = "", y = "AUDPC", title = "Site: Jimma", subtitle = "Severity")+
  coord_flip()
sev_jimma_sk$Means <- round(sev_jimma_sk$Means, 1)

# Means table
p2 + gridExtra::tableGrob(sev_jimma_sk %>% rename(Cultivar = Levels), rows = rownames("Levels"), 
                          theme = ttheme_minimal(padding = unit(c(4, 4.5), "mm")))
```

## Metu

```{r}
sev_metu <- sev %>% 
  filter(site == "metu")
```

### Boxplot

```{r}
sev_metu %>% 
  ggplot(aes(reorder(variety, mean_audpc), mean_audpc))+
  geom_boxplot()
```

### AUDPC vs Max. severity

```{r}
sev_metu %>% 
  ggplot(aes(mean_audpc, mean_max, color = variety))+
  geom_point()
```

##### Mixed model fit

**Fit and residuals**

```{r}
m_audpc <- lmer(mean_audpc ~ variety + (1 | year), data = sev_metu)
simres <- simulateResiduals(m_audpc)
plot(simres)
```

**Obtain estimates of the means and perform multiple comparison**

### Scott-Knott

```{r fig.height=5, fig.width=10}
sev_audpc_metu_sk <- with(sev_metu,
              SK(x= variety,
                 y=mean_audpc,
                 model='y ~ x',
                 which='x'))

plot(sev_audpc_metu_sk)
```

### Mean AUDPC

```{r}
sev_metu_sk <- summary(sev_audpc_metu_sk)
```

### Means and boxplot

```{r fig.height=6.25, fig.width=6}
# severity
p2 <- sev_metu %>% 
  ggplot(aes(reorder(variety, mean_audpc), mean_audpc))+
  geom_boxplot()+
  labs(x = "", y = "AUDPC", title = "Site: Metu", subtitle = "Severity")+
  coord_flip()
sev_metu_sk$Means <- round(sev_metu_sk$Means, 1)

# Means table
p2 + gridExtra::tableGrob(sev_metu_sk %>% rename(Cultivar = Levels), rows = rownames("Levels"), 
                          theme = ttheme_minimal(padding = unit(c(4, 4.5), "mm")))
```

## Tepi

```{r}
sev_tepi <- sev %>% 
  filter(site == "tepi")
```

### Boxplot

```{r}
sev_tepi %>% 
  ggplot(aes(reorder(variety, mean_audpc), mean_audpc))+
  geom_boxplot()
```

### AUDPC vs Max. severity

```{r}
sev_tepi %>% 
  ggplot(aes(mean_audpc, mean_max, color = variety))+
  geom_point()
```

### Mixed model fit

**Fit and residuals**

```{r}
m_audpc <- lmer(mean_audpc ~ variety + (1 | year), data = sev_tepi)
simres <- simulateResiduals(m_audpc)
plot(simres)
```

**Obtain estimates of the means and perform multiple comparison**

### Scott-Knott

```{r}
sev_audpc_tepi_sk <- with(sev_tepi,
              SK(x= variety,
                 y=mean_audpc,
                 model='y ~ x',
                 which='x'))

plot(sev_audpc_tepi_sk)
```

### Means

```{r}
sev_tepi_sk <- summary(sev_audpc_tepi_sk)
```

### Means and boxplot

```{r}
# severity
p2 <- sev_tepi %>% 
  ggplot(aes(reorder(variety, mean_audpc), mean_audpc))+
  geom_boxplot()+
  labs(x = "", y = "AUDPC", title = "Site: tepi", subtitle = "Severity")+
  coord_flip()
sev_tepi_sk$Means <- round(sev_tepi_sk$Means, 1)

# Means table
p2 + gridExtra::tableGrob(sev_tepi_sk %>% rename(Cultivar = Levels), rows = rownames("Levels"), 
                          theme = ttheme_minimal(padding = unit(c(4, 4.5), "mm")))
```

## Mean AUDPC (all sites)

```{r}

sev_agaro_sk %>% 
  bind_rows(sev_gera_sk, sev_jimma_sk, sev_metu_sk, sev_tepi_sk, .id = "Site") %>% 
  remove_rownames() %>% 
  mutate(Site = recode(Site, `1` = "Agaro", `2` = "Gera", `3` = "Jimma", `4` = "Metu", `5` = "Tepi"), 
         Means = round(Means, digits = 1)) %>% 
  unite(Severity, Means, `SK(5%)`, sep = " ", remove = FALSE, na.rm = FALSE) %>%
  dplyr::select(-Means, -`SK(5%)`) %>% 
  spread(key = Site, value = Severity) %>%
  rename(Genotype = Levels) %>% 
  replace(., is.na(.), "-") %>% 
  gt::gt() %>% 
  cols_align(
    align = "right"
  )
  
```









